Peer Assessment 1
---------------
**Jon Rogers**

Let's load the necessary libraries and our data from the working directory

```{r, echo = T}
library(reshape2)
library(ggplot2)
activity <- read.csv("activity.csv")
```

**Total Steps Taken Per Day**

1. Generate a histogram of the total steps taken each day:

```{r, echo = T}
daily.steps <- aggregate(steps ~ date, data = activity, FUN = sum)
barplot(daily.steps$steps, names.arg = daily.steps$date, xlab = "Date", ylab = "Total Steps")
```

2. Report the mean and median total number of steps taken:

The Mean:

```{r, echo = T}
mean(daily.steps$steps)
```
The mean is 10,766 steps

The Median:

```{r, echo = T}
median(daily.steps$steps)
```

The median is 10,765 steps

**Daily Activity Pattern**

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo = T}
interval.steps <- aggregate(steps ~ interval, data = activity, FUN = mean)
plot(interval.steps, type = "l", xlab = "Interval ID", ylab = "Average Steps")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo = T}
interval.steps$interval[which.max(interval.steps$steps)]
```

The interval is #835

**Impute Missing Values**

1. Calculate and report the total number of missing values in the dataset

```{r, echo = T}
sum(is.na(activity))
```

The total number of missing values is 2,304

2. Devise a strategy for filling in all of the missing values in the dataset:

I will fill in the missing data using 0 for the replacement. Yes this is the slacker way to go about it.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in...

```{r, echo = T}
newset <- activity
newset$steps[is.na(newset$steps)] <- 0
```

4A. Make a histogram of the total number of steps taken each day:

```{r, echo = T}
newset.steps <- aggregate(steps ~ date, data = newset, FUN = sum)
barplot(newset.steps$steps, names.arg = newset.steps$date, xlab = "Date", ylab = "Total Steps")
```

Calculate and report the mean and median total number of steps taken:

The Mean:

```{r, echo = T}
mean(newset.steps$steps)
```

The mean is 9,354

The Median:

```{r, echo = T}
median(newset.steps$steps)
```

The median is 10,395

4B. Do these values differ from the estimates from the first part of the assignment? 

Yes. The values are lower becuase of the influence of the zero values replacing the NA's in my strategy. 

4C. What is the impact of imputing missing data on the estimates of the total daily number of steps?

There is an impact on the mean and a small one on the median from the missing data. Using zeros to replace NA's brought both of them down (no surprise) becuase in the inital analysis NA's were discarded.

**Activity Patterns on Weekdays vs. Weekends**

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, echo = T}
daytype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "Weekend"
    } else {
        "Weekday"
    }
}
newset$daytype <- as.factor(sapply(activity$date, daytype))
```

2. Make a panel plot containing a time series plot (i.e. type = "l" ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r, echo = T}
### Alternate code for personal notes (messes with Xaxis)
# newset$interval <- as.factor(newset$interval)
### Daily and Interval Average saved for personal use:
# avg_steps_days = dcast(molten_data, date ~ variable, mean)
# avg_steps_days = dcast(molten_data, interval ~ variable, fun.aggregate = mean)

molten_data = melt(newset, id = c("interval", "date", "daytype"))
avg_steps_days = dcast(molten_data, interval + daytype ~ variable, fun.aggregate = mean)

zPlot <- ggplot(avg_steps_days,aes(x = interval, y = steps, group = daytype, colour = factor(daytype))) 
zPlot + geom_path() + facet_grid(daytype ~ .) + theme(legend.position="none")
```